---
output: github_document
editor_options: 
  chunk_output_type: console
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# zonebuilder

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/zonebuilders/zonebuilder.svg?branch=master)](https://travis-ci.org/zonebuilders/zonebuilder)
<!-- badges: end -->

The goal of zonebuilder is to break up large geographic regions such as cities into manageable zones.
Zoning systems are important in many fields, including transport planning, where we hope to use the zones generated by `zonebuilder` to build the evidence base needed to enable uptake of healthy modes of transport like walking, cycling and renewable electricity-powered public transport. See its website at [zonebuilders.github.io/zonebuilder](https://zonebuilders.github.io/zonebuilder/).

## Installation

<!-- You can install the released version of zonebuilder from [CRAN](https://CRAN.R-project.org) with: -->

```{r, eval=FALSE, echo=FALSE}
install.packages("zonebuilder")
```

Install it from [GitHub](https://github.com/) with:

```{r install-gh, eval=FALSE}
# install.packages("devtools")
devtools::install_github("robinlovelace/zonebuilder")
```

## Using zonebuilder

Zonebuilder works with `sf` objects and works well alongside the `sf` package and visualisation packages that support spatial data such as `ggplot2`, `leaflet`, `mapdeck`, `mapview` and `tmap`, the last of which we'll use in the following maps.
Attaching the package provides the example dataset `zb_region`, the geographic boundary of a large city (London, UK):

```{r}
library(zonebuilder)
library(tmap)
tmap_border = tm_shape(zb_region) + tm_borders()
tmap_border
```

The core functions in the package break up geographical space into discrete chunks.
The syntax is designed to be user friendly.
The following command, for example, breaks London into quadrats of equal size (and keeping only the space inside the boundary):

```{r}
x = zb_region
q = zb_quadrat(x, ncol = 4) # break into 4
qtm(q) 
```

Another zoning option is 'doughnuts':

```{r}
tmap_arrange(
  qtm(zb_doughnut(x, n_circles = 3)) + tmap_border,
  qtm(zb_doughnut(x, distance = 3)) + tmap_border
)
```

A new zoning system implemented in the function `zb_zone()` segments doughnuts of gradually increasing width to create concentric segmented annuli (CSA):

```{r}
library(zonebuilder)
x = zb_region
z = zb_zone(x)
z
qtm(z)
```

This function gives you control over a number of parameters, as documented in the help pages and an in-progress academic paper.

```{r}
args(zb_zone)
```

The zones are labeled by default in the following way. The rings are labels with capital letters, where A is the centre. The segments are numbers using clock hours, so 12, 3, 6, and 9 for north, east, south, west respectivly.  

```{r}
z = zb_zone(x, n_circles = 4)
tm_shape(z) +
  tm_polygons(col = "circle_id", legend.show = FALSE, palette = "magma") + 
  tm_text("label")
```
