---
title: "An automated geographic zoning system for cities with concentric segmented annuli"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An automated geographic zoning system for cities with concentric segmented annuli}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, eval=FALSE, echo=FALSE}
# generate bibliography

citr::tidy_bib_file(rmd_file = "vignettes/paper.Rmd", messy_bibliography = "~/uaf/allrefs.bib", file = "vignettes/references.bib")
```


```{r setup}
library(zonebuilder)
```

Zoning systems have long been a topic of geographic research and political controversy.
From a research perspective, the way in which 

@openshaw_optimal_1977

# The CSA method of zone generation

The 


## Annuli distances

The radius of each annuli in the zoning system can be incremented by a fixed amount, as shown in previous figures.
<!-- Todo: add figures -->
In cases where high geographic resolution is important near the centre of the study region, such as when designing transport systems into the central zone of a city planning, increasing distances between each radius may be desirable.
We experimented with various ways of incrementing the annuli width and suggest linear increases in width as a sensible default for a simple zoning system.
This linear growth leads to distances between each annuli boundary increasing in line with the steps in the [triangular number sequence](https://en.wikipedia.org/wiki/Triangular_number).

# References

